---
title: Working with dplyr and ggplot2
author: ''
date: '2015-12-10'
slug: working-with-dplyr-and-ggplot2
categories:
  - data manipulations
  - data visualizations
tags:
  - dplyr
  - ggplot2
output: 
  html_document: 
    highlight: textmate
    theme: simplex
    toc: yes
---
## dplyr - Data Mainpulation Package

### Intorduction
Most of the data scientists spend 70 to 80% of their daily time on the activities like data preparation which involves data wrangling or cleaning `dplyr` is one of the most popular package which can help them to solve the issues such as **selecting required columns, adding a new column, filtering  usefull observations, even things like sorting or aggregating**    

`dplyr` has couple of functions like  

`select()`  
`filter()`  
`mutate()`  
`arrange()`   
`summarize()`  

and 
**%>% operator**

### load required packages
```{r}
library(dplyr)
library(ggplot2)
```


### loading and examine the dataset
```{r cache=TRUE, collapse=TRUE}
#for illustration purpose take the diamonds dataset from ggplot2 package and attached to this session
data(diamonds)
#examin first 6 observations
head(diamonds)
#take help from r documentation
#?diamonds
#examine the data
dim(diamonds)
str(diamonds)
summary(diamonds)
```
### select()
Select perticular interested columns from the dataset
```{r}
# select only the variables carat, color and price
select(diamonds, carat, color, price)
```

### filter()
filter acts like subsetting the data based on certain conditions
```{r}
# examine the factors in cut variable
table(diamonds$cut)
# subsetting or filtering the diamonds dataset where cut==”Premium”
filter(diamonds, cut=="Premium")
```
### mutate()
Mutate function is generally used to add variables to our dataset
```{r}
diamondsNew<- mutate(diamonds, pricePerCarat = price/carat)
#examine the new dataset whether new variable is added or not
head(diamondsNew)
names(diamondsNew)
```

### arrange()
this function is used to sort or ordering the data
```{r}
# first we will see the first 6 diamonds price in our dataset
head(diamonds$depth)
# then we can use arrange function on top of this vector of first six observationds of depth variable
head(arrange(diamonds,depth))
# the above output is basically shows in ascending order
# you can use desc() function inside arrange to make descending the data
head(arrange(diamonds,desc(depth)))
```

### summarize()
This function is used to get the summary statistics of the data
**its very powerfull when we use this function with the combination of `groupby`**
```{r}
#to get the average of price variable
summarize(diamonds, avgPrice = mean(price, na.rm = TRUE) )
#combination of summarize/summarise with group_by
summarise(group_by(diamonds, cut), mean=mean(price, na.rm = TRUE))
summarize(group_by(diamonds, cut), mean=mean(price, na.rm = TRUE))
```

### %>% operator
The actual power of dplyr package lies in the usage of pipe operator (%>%), its very usefull when ever we required a chain of operations(series of activities) to work on one after another or using one command's result as input for another command   
```{r}
#Now we will use those above functions filter select and mutate and combine them into one and get the result by using %>% operator  
# filter(diamonds, cut=="Premium")
#select(diamonds, carat, color, price)
# diamondsNew<- mutate(diamonds, pricePerCarat = price/carat)


diamondsPipe <- diamonds %>% filter(cut=="Premium") %>% select(carat, color, price) %>% mutate(pricePerCarat = price/carat)
head(diamondsPipe)
```




## ggplot2 - Data Visualization Package

